Probabilistic Model of Bilateral Lymphatic Spread in Head and Neck Cancer

### 2024-08-19 manuscript.qmd
add fig for lateral tumors
Related: #11

The three risk factors for contralateral involvement, midline extension, T-category, and ipsilateral involvement are correlated. For example, midline extension was present in 48% of advanced T-category tumors but only in 7% of early T-category tumors. The higher prevalence of contralateral metastases in advanced T-category tumors may thus be partially explained by a higher rate of midline extension in these tumors. However, figure xxx indicates that advanced T-category and severe ipsilateral involvement do represent additional risk factors . The figure considers only patients with lateralized tumors not extending over the midline. For early T-category and healthy ipsilateral levels I-V, only 1% (1 out of 84 patients) shows involvement of contralateral level II. This increases to 5% (9 out of 165 patients) if ipsilateral level II is involved, to 15% (9 out of 60 patients) if ipsilateral levels II and III are involved, and to 23% (6 out of 26 patients) for advanced T-category with ipsilateral levels II and III involved.

>>>>

The three risk factors for contralateral involvement, midline extension, T-category, and ipsilateral involvement are correlated. For example, midline extension was present in 48% of advanced T-category tumors but only in 7% of early T-category tumors. The higher prevalence of contralateral metastases in advanced T-category tumors may thus be partially explained by a higher rate of midline extension in these tumors. However, @fig-data-strat-uncorr indicates that advanced T-category and severe ipsilateral involvement do represent additional risk factors . The figure considers only patients with lateralized tumors not extending over the midline. For early T-category and healthy ipsilateral levels I-V, only 1% (1 out of 84 patients) shows involvement of contralateral level II. This increases to 5% (9 out of 165 patients) if ipsilateral level II is involved, to 15% (9 out of 60 patients) if ipsilateral levels II and III are involved, and to 23% (6 out of 26 patients) for advanced T-category with ipsilateral levels II and III involved.<!-- TODO: make values here dynamic. -->

![Contralateral involvement prevalence by LNL for lateralized tumors. Shown are four different scenarios: 1) early T-category with an ipsilateral N0 neck (green), 2) early T-category with ipsilaterally LNL II involved (blue), 3) late T-category with ipsilateral LNL II metastatic (orange), and 4) early T-category with ipsilateral both levels II and III involved (red). This illustrates that even when controlling for the tumor's lateralization, T-category and ipsilaeral involvement are predictive of contralateral metastases.](figures/fig_data_strat_uncorr.svg){#fig-data-strat-uncorr}
